Component response rate variation underlies the stability of complex systems

### 2020-03-20 notebook/SI.Rmd
Adjustments to the notation of the SI to match the main text

**This supplemental information supports the manuscript "Component response rate variation drives stability in large complex systems" with additional analyses to support its conclusions. All text, code, and data underlying this manuscript are publicly available on [GitHub](https://github.com/bradduthie/RandomMatrixStability) as part of the RandomMatrixStability R package.**

@@

- [Stability across increasing $S$](#IncrS) 
- [Stability of ecological networks](#ecological) 
    - [Competitor networks](#competition) 
    - [Mutualist networks](#mutualism) 
    - [Predator-prey networks](#pred-prey) 
- [Sensitivity of connectance (C) values](#connectance)
    - [C = 0.3](#connect3)
    - [C = 0.5](#connect5)
    - [C = 0.7](#connect7)
    - [C = 0.9](#connect9)
- [Large networks of $C = 0.05$ across $S$ and $\sigma$](#sigma)  
    - [$\sigma$ = 0.3](#sigma3)
    - [$\sigma$ = 0.4](#sigma4)
    - [$\sigma$ = 0.5](#sigma5)
    - [$\sigma$ = 0.6](#sigma6)
- [Sensitivity of distribution of $\gamma$](#gam_dist)
- [Stability of structured networks](#structured)
- [Feasibility of complex systems](#Feasibility)
- [Stability given targeted manipulation of $\gamma$ (genetic algorithm)](#ga)
- [Consistency with Gibbs et al. (2018)](#Gibbs)

@@

Figure 4 of the main text reports the number of stable random complex systems found over 1 million iterations. The table below shows the results for all simulations of random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ given a range of $S = \{2, 3, ..., 49, 50\}$. In this table, the `A` refers to $\mathbf{A}$ matrices where $\gamma = 1$, while `M` refers to $\mathbf{M}$ matrices after $\sigma^{2}_{\gamma}$ is added and $\gamma \sim \mathcal{U}(0, 2)$. Each row summarises data for a given $S$ over 1 million randomly simulated $\mathbf{M}$. The column `A_unstable` shows the number of $\mathbf{A}$ matrices that are unstable, and the column `A_stable` shows the number of $\mathbf{A}$ matrices that are stable (these two columns sum to 1 million). Similarly, the column `M_unstable` shows the number of $\mathbf{M}$ matrices that are unstable and `M_stable` shows the number that are stable. The columns `A_stabilised` and `A_destabilised` show how many $\mathbf{M}$ matrices were stabilised or destabilised, respectively, by $\sigma^{2}_{\gamma}$.

@@

Stability of ecological networks {#ecological}
================================================================================

@@

In the main text, for simplicity, I assumed connectance values of $C = 1$, meaning that all off-diagonal elements of a matrix $\mathbf{M}$ were potentially nonzero and sampled from a normal distribution $\mathcal{N}(0, \sigma^{2})$ where $\sigma = 0.4$. Here I present four tables showing the number of stable communities given $C = \{0.3, 0. 5, 0.7, 0.9 \}$. In all cases, uniform variation in component response rate ($\gamma \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma = 1$). In contrast to the main text, 100000 rather than 1 million $\mathbf{M}$ were simulated. As with the results on [stability with increasing $S$](#IncrS) shown above, in the tables below `A` refers to $\mathbf{A}$ matrices when $\gamma = 1$, and `M` refers to $\mathbf{M}$ matrices after $\sigma^{2}_{\gamma}$ is added. The column `A_unstable` shows the number of $\mathbf{A}$ matrices that are unstable, and the column `A_stable` shows the number of $\mathbf{A}$ matrices that are stable (these two columns sum to 100000). Similarly, the column `M_unstable` shows the number of $\mathbf{M}$ matrices that are unstable and `M_stable` shows the number that are stable. The columns `A_stabilised` and `A_destabilised` show how many $\mathbf{A}$ matrices were stabilised or destabilised, respectively, by $\sigma^{2}_{\gamma}$.

@@

Sensitivity of interaction strength ($\sigma$) values {#sigma}  
================================================================================

Results below show stability results given varying interaction strengths ($\sigma$) for $C = 0.05$ (note that system size $S$ values are larger and increase by 10 with increasing rows). In the tables below (as [above](#IncrS)), `A` and `M` refers to matrices for $\gamma = 1$ and $\sigma^{2}_{\gamma}$, respectively.

@@

<a name="sigma3">**Interaction strength $\mathbf{\sigma = 0.3}$**</a>

@@

<a name="sigma4">**Interaction strength $\mathbf{\sigma = 0.4}$**</a>

@@

<a name="sigma5">**Interaction strength $\mathbf{\sigma = 0.5}$**</a>

@@

<a name="sigma6">**Interaction strength $\mathbf{\sigma = 0.6}$**</a>

>>>>

**This supplemental information supports the manuscript "Component response rate variation underlies the stability of complex systems" with additional analyses to support its conclusions. All text, code, and data underlying this manuscript are publicly available on [GitHub](https://github.com/bradduthie/RandomMatrixStability) as part of the RandomMatrixStability R package.**

@@

- [Stability across increasing $S$](#IncrS) 
- [Stability of random ecological networks](#ecological) 
    - [Competitor networks](#competition) 
    - [Mutualist networks](#mutualism) 
    - [Predator-prey networks](#pred-prey) 
- [Sensitivity of connectance (C) values](#connectance)
    - [C = 0.3](#connect3)
    - [C = 0.5](#connect5)
    - [C = 0.7](#connect7)
    - [C = 0.9](#connect9)
- [Large networks of $C = 0.05$ across $S$ and $\sigma$](#sigma)  
    - [$\sigma$ = 0.3](#sigma3)
    - [$\sigma$ = 0.4](#sigma4)
    - [$\sigma$ = 0.5](#sigma5)
    - [$\sigma$ = 0.6](#sigma6)
- [Sensitivity of distribution of $\gamma$](#gam_dist)
- [Stability of structured networks](#structured)
- [Feasibility of complex systems](#Feasibility)
- [Stability given targeted manipulation of $\gamma$ (genetic algorithm)](#ga)
- [Consistency with Gibbs et al. (2018)](#Gibbs)

@@

Figure 4 of the main text reports the number of stable random complex systems found over 1 million iterations. The table below shows the results for all simulations of random $\mathbf{M}$ matrices at $\sigma_{A} = 0.4$ and $C = 1$ given a range of $S = \{2, 3, ..., 49, 50\}$. In this table, the `A` refers to $\mathbf{A}$ matrices where $\gamma = 1$, while `M` refers to $\mathbf{M}$ matrices after $\sigma^{2}_{\gamma}$ is added and $\gamma \sim \mathcal{U}(0, 2)$. Each row summarises data for a given $S$ over 1 million randomly simulated $\mathbf{M}$. The column `A_unstable` shows the number of $\mathbf{A}$ matrices that are unstable, and the column `A_stable` shows the number of $\mathbf{A}$ matrices that are stable (these two columns sum to 1 million). Similarly, the column `M_unstable` shows the number of $\mathbf{M}$ matrices that are unstable and `M_stable` shows the number that are stable. The columns `A_stabilised` and `A_destabilised` show how many $\mathbf{M}$ matrices were stabilised or destabilised, respectively, by $\sigma^{2}_{\gamma}$.

@@

Stability of random ecological networks {#ecological}
================================================================================

@@

In the main text, for simplicity, I assumed connectance values of $C = 1$, meaning that all off-diagonal elements of a matrix $\mathbf{M}$ were potentially nonzero and sampled from a normal distribution $\mathcal{N}(0, \sigma^{2}_{A})$ where $\sigma_{A} = 0.4$. Here I present four tables showing the number of stable communities given $C = \{0.3, 0. 5, 0.7, 0.9 \}$. In all cases, uniform variation in component response rate ($\gamma_{i} \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma_{i} = 1$). In contrast to the main text, 100000 rather than 1 million $\mathbf{M}$ were simulated. As with the results on [stability with increasing $S$](#IncrS) shown above, in the tables below `A` refers to $\mathbf{A}$ matrices when $\gamma = 1$, and `M` refers to $\mathbf{M}$ matrices after $\sigma^{2}_{\gamma}$ is added. The column `A_unstable` shows the number of $\mathbf{A}$ matrices that are unstable, and the column `A_stable` shows the number of $\mathbf{A}$ matrices that are stable (these two columns sum to 100000). Similarly, the column `M_unstable` shows the number of $\mathbf{M}$ matrices that are unstable and `M_stable` shows the number that are stable. The columns `A_stabilised` and `A_destabilised` show how many $\mathbf{A}$ matrices were stabilised or destabilised, respectively, by $\sigma^{2}_{\gamma}$.

@@

Sensitivity of interaction strength ($\sigma_{A}$) values {#sigma}  
================================================================================

Results below show stability results given varying interaction strengths ($\sigma_{A}$) for $C = 0.05$ (note that system size $S$ values are larger and increase by 10 with increasing rows). In the tables below (as [above](#IncrS)), `A` and `M` refers to matrices for $\gamma = 1$ and $\sigma^{2}_{\gamma}$, respectively.

@@

<a name="sigma3">**Interaction strength $\mathbf{\sigma_{A} = 0.3}$**</a>

@@

<a name="sigma4">**Interaction strength $\mathbf{\sigma_{A} = 0.4}$**</a>

@@

<a name="sigma5">**Interaction strength $\mathbf{\sigma_{A} = 0.5}$**</a>

@@

<a name="sigma6">**Interaction strength $\mathbf{\sigma_{A} = 0.6}$**</a>

### 2020-03-20 notebook/SI.tex
Adjustments to the notation of the SI to match the main text

\textbf{This supplemental information supports the manuscript
``Component response rate variation drives stability in large complex
systems'' with additional analyses to support its conclusions. All text,
code, and data underlying this manuscript are publicly available on
\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub} as
part of the RandomMatrixStability R package.}

@@

\begin{itemize}
\tightlist
\item
  \protect\hyperlink{IncrS}{Stability across increasing \(S\)}
\item
  \protect\hyperlink{ecological}{Stability of ecological networks}

@@

Figure 4 of the main text reports the number of stable random complex
systems found over 1 million iterations. The table below shows the
results for all simulations of random \(\mathbf{M}\) matrices at
\(\sigma = 0.4\) and \(C = 1\) given a range of
\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A} refers to
\(\mathbf{A}\) matrices where \(\gamma = 1\), while \texttt{M} refers to
\(\mathbf{M}\) matrices after \(\sigma^{2}_{\gamma}\) is added and
\(\gamma \sim \mathcal{U}(0, 2)\). Each row summarises data for a given
\(S\) over 1 million randomly simulated \(\mathbf{M}\). The column
\texttt{A\_unstable} shows the number of \(\mathbf{A}\) matrices that
are unstable, and the column \texttt{A\_stable} shows the number of
\(\mathbf{A}\) matrices that are stable (these two columns sum to 1
million). Similarly, the column \texttt{M\_unstable} shows the number of
\(\mathbf{M}\) matrices that are unstable and \texttt{M\_stable} shows
the number that are stable. The columns \texttt{A\_stabilised} and
\texttt{A\_destabilised} show how many \(\mathbf{M}\) matrices were
stabilised or destabilised, respectively, by \(\sigma^{2}_{\gamma}\).

@@

\hypertarget{ecological}{\section{Stability of ecological
networks}\label{ecological}}

@@

In the main text, for simplicity, I assumed connectance values of
\(C = 1\), meaning that all off-diagonal elements of a matrix
\(\mathbf{M}\) were potentially nonzero and sampled from a normal
distribution \(\mathcal{N}(0, \sigma^{2})\) where \(\sigma = 0.4\). Here
I present four tables showing the number of stable communities given
\(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases, uniform variation in
component response rate (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a
higher number of stable communities than when \(\gamma\) did not vary
(\(\gamma = 1\)). In contrast to the main text, 100000 rather than 1
million \(\mathbf{M}\) were simulated. As with the results on
\protect\hyperlink{IncrS}{stability with increasing \(S\)} shown above,
in the tables below \texttt{A} refers to \(\mathbf{A}\) matrices when
\(\gamma = 1\), and \texttt{M} refers to \(\mathbf{M}\) matrices after
\(\sigma^{2}_{\gamma}\) is added. The column \texttt{A\_unstable} shows
the number of \(\mathbf{A}\) matrices that are unstable, and the column
\texttt{A\_stable} shows the number of \(\mathbf{A}\) matrices that are
stable (these two columns sum to 100000). Similarly, the column
\texttt{M\_unstable} shows the number of \(\mathbf{M}\) matrices that
are unstable and \texttt{M\_stable} shows the number that are stable.
The columns \texttt{A\_stabilised} and \texttt{A\_destabilised} show how
many \(\mathbf{A}\) matrices were stabilised or destabilised,
respectively, by \(\sigma^{2}_{\gamma}\).

@@

\hypertarget{sigma}{\section{\texorpdfstring{Sensitivity of interaction
strength (\(\sigma\))
values}{Sensitivity of interaction strength (\textbackslash{}sigma) values}}\label{sigma}}

Results below show stability results given varying interaction strengths
(\(\sigma\)) for \(C = 0.05\) (note that system size \(S\) values are
larger and increase by 10 with increasing rows). In the tables below (as
\protect\hyperlink{IncrS}{above}), \texttt{A} and \texttt{M} refers to
matrices for \(\gamma = 1\) and \(\sigma^{2}_{\gamma}\), respectively.

\textbf{Interaction strength \(\mathbf{\sigma = 0.3}\)}

@@

\textbf{Interaction strength \(\mathbf{\sigma = 0.4}\)}

@@

\textbf{Interaction strength \(\mathbf{\sigma = 0.5}\)}

@@

\textbf{Interaction strength \(\mathbf{\sigma = 0.6}\)}

>>>>

\textbf{This supplemental information supports the manuscript
``Component response rate variation underlies the stability of complex
systems'' with additional analyses to support its conclusions. All text,
code, and data underlying this manuscript are publicly available on
\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub} as
part of the RandomMatrixStability R package.}

@@

\begin{itemize}
\tightlist
\item
  \protect\hyperlink{IncrS}{Stability across increasing \(S\)}
\item
  \protect\hyperlink{ecological}{Stability of random ecological
  networks}

@@

Figure 4 of the main text reports the number of stable random complex
systems found over 1 million iterations. The table below shows the
results for all simulations of random \(\mathbf{M}\) matrices at
\(\sigma_{A} = 0.4\) and \(C = 1\) given a range of
\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A} refers to
\(\mathbf{A}\) matrices where \(\gamma = 1\), while \texttt{M} refers to
\(\mathbf{M}\) matrices after \(\sigma^{2}_{\gamma}\) is added and
\(\gamma \sim \mathcal{U}(0, 2)\). Each row summarises data for a given
\(S\) over 1 million randomly simulated \(\mathbf{M}\). The column
\texttt{A\_unstable} shows the number of \(\mathbf{A}\) matrices that
are unstable, and the column \texttt{A\_stable} shows the number of
\(\mathbf{A}\) matrices that are stable (these two columns sum to 1
million). Similarly, the column \texttt{M\_unstable} shows the number of
\(\mathbf{M}\) matrices that are unstable and \texttt{M\_stable} shows
the number that are stable. The columns \texttt{A\_stabilised} and
\texttt{A\_destabilised} show how many \(\mathbf{M}\) matrices were
stabilised or destabilised, respectively, by \(\sigma^{2}_{\gamma}\).

@@

\hypertarget{ecological}{\section{Stability of random ecological
networks}\label{ecological}}

@@

In the main text, for simplicity, I assumed connectance values of
\(C = 1\), meaning that all off-diagonal elements of a matrix
\(\mathbf{M}\) were potentially nonzero and sampled from a normal
distribution \(\mathcal{N}(0, \sigma^{2}_{A})\) where
\(\sigma_{A} = 0.4\). Here I present four tables showing the number of
stable communities given \(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases,
uniform variation in component response rate
(\(\gamma_{i} \sim \mathcal{U}(0, 2)\)) led to a higher number of stable
communities than when \(\gamma\) did not vary (\(\gamma_{i} = 1\)). In
contrast to the main text, 100000 rather than 1 million \(\mathbf{M}\)
were simulated. As with the results on
\protect\hyperlink{IncrS}{stability with increasing \(S\)} shown above,
in the tables below \texttt{A} refers to \(\mathbf{A}\) matrices when
\(\gamma = 1\), and \texttt{M} refers to \(\mathbf{M}\) matrices after
\(\sigma^{2}_{\gamma}\) is added. The column \texttt{A\_unstable} shows
the number of \(\mathbf{A}\) matrices that are unstable, and the column
\texttt{A\_stable} shows the number of \(\mathbf{A}\) matrices that are
stable (these two columns sum to 100000). Similarly, the column
\texttt{M\_unstable} shows the number of \(\mathbf{M}\) matrices that
are unstable and \texttt{M\_stable} shows the number that are stable.
The columns \texttt{A\_stabilised} and \texttt{A\_destabilised} show how
many \(\mathbf{A}\) matrices were stabilised or destabilised,
respectively, by \(\sigma^{2}_{\gamma}\).

@@

\hypertarget{sigma}{\section{\texorpdfstring{Sensitivity of interaction
strength (\(\sigma_{A}\))
values}{Sensitivity of interaction strength (\textbackslash{}sigma\_\{A\}) values}}\label{sigma}}

Results below show stability results given varying interaction strengths
(\(\sigma_{A}\)) for \(C = 0.05\) (note that system size \(S\) values
are larger and increase by 10 with increasing rows). In the tables below
(as \protect\hyperlink{IncrS}{above}), \texttt{A} and \texttt{M} refers
to matrices for \(\gamma = 1\) and \(\sigma^{2}_{\gamma}\),
respectively.

\textbf{Interaction strength \(\mathbf{\sigma_{A} = 0.3}\)}

@@

\textbf{Interaction strength \(\mathbf{\sigma_{A} = 0.4}\)}

@@

\textbf{Interaction strength \(\mathbf{\sigma_{A} = 0.5}\)}

@@

\textbf{Interaction strength \(\mathbf{\sigma_{A} = 0.6}\)}
